ENME 625 \\ Multi-Objective Optimization with Genetic Algorithms

### 2017-05-09 draft1_SM.tex
Update draft1_SM.tex

This report develops multi-objective optimization methods using genetic algorithms. Three cases of multi-objective optimization are considered: unconstrained optimization, constrained optimization, and robust optimization. A fitness function for the first two cases was developed using the built-in MATLAB 'ga' routine. The last case was developed such that uncertainty in function parameters could be handled and robustly optimized against.

@@

\noindent Within NSGA, there are certain parameters which affect the final generation of the Pareto frontier. It was found for the test problems, the selection of the fitness degradation parameter, $\epsilon$, and the sharing parameter, $\sigma$, were the most important. The fitness degradation parameter, $\epsilon$, affects the difference in initial fitness values between layers. The larger the value of $\epsilon$, the higher a penalty is placed on individuals in non-dominated layers that are not in the Pareto layer. The sharing parameter, $\sigma$, affects the closeness of the final Pareto points. The higher the value of $\sigma$, the more "gaps" in the frontier.  \newline

@@

\noindent The true Pareto frontier for this problem occurs when $x_i = 0$ for i = 2,...,30. Figure~\ref{fig:ZDT1} shows a sample result from both MATLABs built in MOGA and the MOGA developed in this project. The quality metrics chosen to evaluate this problem are Coverage Difference (CD)and Pareto Spread (OS). Ten runs for each algorithm were performed and the mean and standard deviation of each metric are tabulated in Table~\ref{tab:ZDT1}.

@@

\noindent From these metrics, there is certainly a trade-off between MATLABs MOGA and the MOGA developed in this project. The coverage difference of the new MOGA is better in this problem whereas the Pareto spread is improved when using MATLABs MOGA.

@@

\noindent The true Pareto frontier for this problem, similar to ZDT1, occurs when $x_i$ = 0 for i = 2,...,30. Figure~\ref{fig:ZDT2} shows a sample result from both MATLABs built in MOGA and the MOGA developed in this project. Table~\ref{tab:ZDT2} summarizes the mean quality metrics for each algorithm for ten runs.
\begin{figure}[H]
  \caption{Example Pareto Results for ZDT2}
  \centering
  \includegraphics[width=0.85\textwidth]{ZDT2_pareto_final.png}  
  \label{fig:ZDT2}
\end{figure}

@@

\noindent The true Pareto frontier for this problem again occurs when $x_i$ = 0 for i = 2,...,30. Figure~\ref{fig:ZDT3} shows a sample result from both MATLABs built in MOGA and the MOGA developed in this project. Table~\ref{tab:ZDT3} summarizes the mean quality metrics for each algorithm for ten runs. \newline

@@

\noindent Overall, our MOGA outperformed MATLABs MOGA in both quality metrics. However, by examining the Pareto frontiers, it is clear that neither algorithm is truely satisfactory in estimating the true Pareto frontier.

@@

There are many ways in which fitness evaluation can be accomplished for constraints. In general, some factors which should be included in fitness evaluation are the feasibility of an individual, the amount of infeasibility, and the number of violated constraints for a certain individual. Several methods to take into account these factors are described in \cite{kurpati2002constraint}. For this project, all three parameters are used to evaluated the fitness of individuals in constrained problems. This method is denoted at CH-I4 in \cite{kurpati2002constraint}. \newline

@@

\subsection{OSY} 
This test problem, denoted as OSY in \cite{deb2001multi} is shown below, Figure~\ref{fig:OSY_true} shows the true Pareto frontier for this problem.

@@

\begin{table}[H]
\caption{Quality Metrics for OSY} 
\centering 
\begin{tabular}{|c|c|c|} 
\hline\hline  
Metric & MATLABs MOGA & Our MOGA \\ \hline
CD & 0.7882 (0.1995) & 0.5507 (0.0685)  \\ \hline
OS & 0.5322 (0.3393) & 0.9608 (0.2427) \\ \hline
\end{tabular}
\label{tab:OSY} 
\end{table}
\begin{figure}[H]
  \caption{OSY Quality Metrics}
  \centering
  \includegraphics[width=0.85\textwidth]{OSY_QM.png}  
  \label{fig:OSY_QM}
\end{figure}
\newpage
\subsection{TNK} 
This test problem, denoted as TNK in \cite{deb2001multi} is shown below with the true Pareto frontier shown in Figure~\ref{fig:TNK_true}.

@@

\noindent Comparing the estimated Pareto frontiers to the true Pareto frontier, is is clear that our MOGA outperforms MATLAB's MOGA. The Pareto spread in our MOGA is significantly higher however the coverage difference is higher in MATLABs MOGA (p<0.05).
\newpage
\subsection{CTP} 
This test problem, denoted as CTP in \cite{deb2001multi} is shown below and the true Pareto frontier is shown in Figure~\ref{fig:CTP_true}.

@@

\noindent Figure~\ref{fig:robust} shows a sample result from both MATLABs built in MOGA and the robust MOGA developed in this project. Table~\ref{tab:robustTNK} summarizes the mean quality metrics for each algorithm for ten runs.
\begin{figure}[H]
  \caption{Example Pareto Results for TNK}
  \centering
  \includegraphics[width=0.85\textwidth]{prob7_nChr25_nRun400.png}  
  \label{fig:robustTNK}
\end{figure}

@@

>>>>

This report develops multi-objective optimization methods using genetic algorithms. Three cases of multi-objective optimization are considered: unconstrained optimization, constrained optimization, and robust optimization. A fitness function for the first two cases was developed using the built-in MATLAB 'ga' routine. The last case was developed such that uncertainty in function parameters could be handled and robustly optimized against. \newline

@@

\noindent Within NSGA, there are certain parameters which affect the final generation of the Pareto frontier. It was found for the test problems, the selection of the fitness degradation parameter, $\epsilon$, and the sharing parameter, $\sigma$, were the most important. The fitness degradation parameter, $\epsilon$, affects the difference in initial fitness values between layers. The larger the value of $\epsilon$, the higher a penalty is placed on individuals in non-dominated layers that are not in the Pareto layer. The sharing parameter, $\sigma$, affects the closeness of the final Pareto points. The higher the value of $\sigma$, the more "gaps" in the frontier. For each problem, these parameters were randomly tuned such that the "best" Pareto frontier (in terms of coverage difference and Pareto spread). \newline

@@

\noindent The true Pareto frontier for this problem occurs when $x_i = 0$ for i = 2,...,30. Figure~\ref{fig:ZDT1} shows a sample result from both MATLABs built in MOGA and the MOGA developed in this project. In this problem, $\alpha=1$, $\sigma = 0.158$, and $\epsilon = 0.22$. The final results were found using 300 chromosomes in each population. The quality metrics chosen to evaluate this problem (and all other problems) are Coverage Difference (CD) and Pareto Spread (OS). Ten runs for each algorithm were performed using the same initialization parameters and the mean and standard deviation of each metric are tabulated in Table~\ref{tab:ZDT1}.

@@

\noindent From these metrics, there is certainly a trade-off between MATLABs MOGA and the MOGA developed in this project. The coverage difference of the new MOGA is lower (and therefore superior) in this problem whereas the Pareto spread is improved when using MATLABs MOGA.

@@

\noindent The true Pareto frontier for this problem, similar to ZDT1, occurs when $x_i$ = 0 for i = 2,...,30. Selected parameters were similar to ZDT1 where $\alpha = 0.5$, $\sigma=0.5$, $\epsilon=0.22$; the number of chromosomes was 300. Figure~\ref{fig:ZDT2} shows a sample result from both MATLABs built in MOGA and the MOGA developed in this project. Table~\ref{tab:ZDT2} summarizes the mean quality metrics for each algorithm over ten runs.
\begin{figure}[H]
  \caption{Example Pareto Results for ZDT2}
  \centering
  \includegraphics[width=0.85\textwidth]{ZDT2_pareto_final.png}  
  \label{fig:ZDT2}
\end{figure}

@@

\noindent The true Pareto frontier for this problem again occurs when $x_i$ = 0 for i = 2,...,30. The fitness parameters selected in this problem were $\alpa = 1$, $\sigma = 0158$ and $\epsilon = 0.22$. Again, the number of chromosomes in this problem was 300. Figure~\ref{fig:ZDT3} shows a sample result from both MATLABs built in MOGA and the MOGA developed in this project. Table~\ref{tab:ZDT3} summarizes the mean quality metrics for each algorithm for ten runs. \newline

@@

@@

There are many ways in which fitness evaluation can be accomplished for constrainted problems. In general, some factors which may be included in fitness evaluation are the feasibility of an individual, the amount of infeasibility, and the number of violated constraints for a certain individual. Several methods to take into account these factors are described in \cite{kurpati2002constraint}. For this project, all three parameters are used to evaluated the fitness of individuals in constrained problems. This method is denoted at CH-I4 in \cite{kurpati2002constraint}. \newline

@@

\subsection{OSY} 
This test problem, denoted as OSY in \cite{deb2001multi} is shown below, Figure~\ref{fig:OSY_true} shows the true Pareto frontier for this problem. The factors chosen for this problem were $CF1 = 0.005$ and $CF2 = 0.015$, this placed a higher weight of the number of constraint violations than the amount of infeasibility.

@@

\begin{table}[H]
\caption{Quality Metrics for OSY} 
\centering 
\begin{tabular}{|c|c|c|} 
\hline\hline  
Metric & MATLABs MOGA & Our MOGA \\ \hline
CD & 0.7882 (0.1995) & 0.5507 (0.0685)  \\ \hline
OS & 0.5322 (0.3393) & 0.9608 (0.2427) \\ \hline
\end{tabular}
\label{tab:OSY} 
\end{table}
\begin{figure}[H]
  \caption{OSY Quality Metrics}
  \centering
  \includegraphics[width=0.85\textwidth]{OSY_QM.png}  
  \label{fig:OSY_QM}
\end{figure}
\newpage
\subsection{TNK} 
This test problem, denoted as TNK in \cite{deb2001multi} is shown below with the true Pareto frontier shown in Figure~\ref{fig:TNK_true}. Constraint parameters CF1 and CF2 were both chosen to be 0.01 to place equal emphasis on amount of infeasibility and number of violated constraints.

@@

\noindent Comparing the estimated Pareto frontiers to the true Pareto frontier, is is clear that our MOGA outperforms MATLAB's MOGA. The Pareto spread in our MOGA is significantly higher however the coverage difference is higher in MATLABs MOGA (p<0.05).
\newpage
\subsection{CTP} 
This test problem, denoted as CTP in \cite{deb2001multi} is shown below and the true Pareto frontier is shown in Figure~\ref{fig:CTP_true}. Just as in TNK, constraint parameters CF1 and CF2 were both chosen to be 0.01 to place equal emphasis on amount of infeasibility and number of violated constraints.

@@

\noindent Figure~\ref{fig:robust} shows a sample result from both MATLABs built in MOGA and the robust MOGA developed in this project. Table~\ref{tab:robustTNK} summarizes the mean quality metrics for each algorithm for ten runs. \newline
\begin{figure}[H]
  \caption{Example Pareto Results for TNK}
  \centering
  \includegraphics[width=0.85\textwidth]{prob7_nChr25_nRun400.png}  
  \label{fig:robustTNK}
\end{figure}

@@

\section{Conclusion}

Overall, it was found that different optimization algorithms result in a range of solutions for a given test problem. Thus it is important, when developing an optimization algorithm, to test against a range of optimization problems. In addition, it was also concluded that special attention should be paid to tunable parameters within each algorithm. Understanding the nature of the parameter and how it affects the algorithm is crucial to obtaining the "true" Pareto frontier.
